---
title: "SMARTE_FINAL"
output: html_document
---


#### Load packages

```{r}
library(googledrive)
library(psych) # for descriptives
library(tidyverse)
library(ggplot2)
library(tidyr)
library(dplyr)
library(scales)
library(knitr)
library(tidytext)
library(corrplot)
library(FactoMineR)
library(factoextra)
library(gridExtra)
library(mclust)
library(ggrepel)
library(ranger)
```

#### Access google drive to get files
```{r}
drive_auth()
```

```{r}
# ID of parent folder
drive_ls(path = as_id("1szhWdy3lHwFRDwwxv8IgF1NDq5fYn43I"))

# Grab JV's merged Study3 x school data
drive_download(file = as_id("1Gt7K-XIkHKzSjW-VxB2cvQyrL25hLZ-z"), path = "SMARTE_merged.csv", overwrite = TRUE)
```

```{r}
JV_merged <- read.csv("SMARTE_merged.csv")
```

```{r}
#Grabbing School Data

drive_ls(path = as_id("1szhWdy3lHwFRDwwxv8IgF1NDq5fYn43I"))
drive_download(file = as_id("12xVJ47voWUd3gI2VCdDRm-slzGGosnTA"), path = "SchoolData_1_3.csv", overwrite = TRUE)
```

```{r}
school <- read.csv("SchoolData_1_3.csv")
```

```{r}
# Study 3 data
Study3 <- read_csv("SMARTE_Brazil3.csv")
```

```{r}
# Grabbing Study 1 data
drive_ls(path = as_id("1szhWdy3lHwFRDwwxv8IgF1NDq5fYn43I"))
drive_download(file = as_id("105BRxj2sDnm56zZZRuk6n7rtLN91TRdt"), path = "Brazil_Study1.csv", overwrite = TRUE)
```

```{r}
Study1 <- read.csv("Brazil_Study1.csv")
```


#### Merge Study 1 and school data
```{r}
# Making sure PIDs are characters to prevent merge errors
Study1$pid <- as.character(Study1$pid)
school$pid <- as.character(school$pid)

# Merging to new dataframe
Study1_X_school <- Study1 %>%
  inner_join(school, by = "pid")

# Checking to see total sample size
length(unique(Study1_X_school$pid))
```


### Descriptive Analysis
#### Looking into phases
```{r}
table(Study1_X_school$phase, useNA = "ifany")
```

#### Checking to see if NA is PhaseA (Enumeration), and it was
```{r}
Study1_X_school %>%
  filter(is.na(phase)) %>%
  select(Type, module) %>%
  table()
```

#### Updating the phase names
```{r}
Study1_X_school <- Study1_X_school %>%
  mutate(phase = case_when(
    is.na(phase)    ~ "A",
    phase == "TestA" ~ "B",
    phase == "TestB" ~ "C",
    phase == "TestC" ~ "D",
    phase == "TestD" ~ "E",
    TRUE            ~ as.character(phase)
  ))
```


#### Finding optimal RT per phase
```{r}
# Creating new data frame to use for optimal reaction time per phase
plot_data <- Study1_X_school %>%
  filter(!is.na(Submit_RT), !is.na(correct)) %>%
  mutate(Submit_RT_sec = Submit_RT / 1000) %>% #Converting into seconds
  group_by(phase) %>%
  mutate(Time_Bucket = ntile(Submit_RT_sec, 15)) %>%
  group_by(phase, Time_Bucket) %>%
  summarize(
    Average_RT = mean(Submit_RT_sec, na.rm = TRUE),
    Accuracy = mean(correct, na.rm = TRUE),
    .groups = "drop"
  )

# Create the optimal_lines data frame
optimal_lines <- plot_data %>%
  group_by(phase) %>%
  slice_max(order_by = Accuracy, n = 1, with_ties = FALSE)

# Plot
ggplot(data = plot_data, aes(x = Average_RT, y = Accuracy)) +
  geom_point(color = "#5DA5DA", size = 3, alpha = 0.8) +
  geom_smooth(method = "loess", color = "mediumblue", se = FALSE) +
  geom_vline(data = optimal_lines, aes(xintercept = Average_RT),
             color = "red", linetype = "dashed", alpha = 0.6) +
  geom_text(data = optimal_lines, 
            aes(x = Average_RT, y = min(plot_data$Accuracy), 
                label = paste("Optimal RT ≈", round(Average_RT, 2))),
            color = "red", angle = 90, vjust = -1, hjust = 0) +
  facet_wrap(~ phase, scales = "free_x") +
  labs(title = "Speed-Accuracy Curves by Phase (Study 1)",
       x = "Average Response Time (Seconds)",
       y = "Accuracy") +
  theme_bw()
```


```{r}
# Formatting optimal_lines into a clean table we can look up later if we need to
opt_lookup <- optimal_lines %>%
  select(phase, Optimal_RT_sec = Average_RT) #Keeping only the phase and the optimal response time

# Joining it to Study1_x_school dataset
Study1_X_school <- Study1_X_school %>%
  left_join(opt_lookup, by = "phase") %>% # Attaching the optimal times based on the phase

# Calculating the distance from optimal (in seconds)  
  mutate(
    Submit_RT_sec = Submit_RT / 1000, # First converting the time to seconds
    Distance_From_Optimal = abs(Submit_RT_sec - Optimal_RT_sec) # Then, calculating the absolute distance
  )

# Checking to see if it worked
Study1_X_school %>%
  select(phase, Submit_RT_sec, Optimal_RT_sec, Distance_From_Optimal) %>%
  head()
```



### Looking at which school and characteristic variables could effect students to deviate from optimal response time by random forest

```{r}
# Preparing data for random forest
rf_data_context <- Study1_X_school %>%
  filter(phase != "E") %>% # Dropping Phase E since N=7 is too small
  select(Distance_From_Optimal, GRADE, SES10, GENDER, Nervous_Level_Num) %>% # Selecting only the target and the predictors
  na.omit() # Removing rows with missing data in these specific columns

# Running the Random Forest
set.seed(123) #for reproducability
model_a_context <- ranger(
  formula = Distance_From_Optimal ~ ., 
  data = rf_data_context, 
  importance = "permutation",
  num.trees = 500
)

# Printing results
print(sort(model_a_context$variable.importance, decreasing = TRUE))
```


### Visualizing importance scores
```{r}
# New dataframe 
imp_model_a <- data.frame(
  Feature = names(model_a_context$variable.importance),
  Importance = as.numeric(model_a_context$variable.importance)
)

ggplot(imp_model_a, aes(x = reorder(Feature, Importance), y = Importance)) +
  geom_col(fill = "#5DA5DA", width = 0.7) +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Feature importance scores", 
    x = "Features", 
    y = "Permutation Importance Score"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 12)
  )
```

### Deeper analysis, looking into each feature
#### GRADE first
```{r}
# Boxplots
ggplot(Study1_X_school %>% 
         filter(phase != "E"),
       aes(x = as.factor(GRADE), y = Distance_From_Optimal, fill = as.factor(GRADE))) + #Treating GRADE as distinct categories rather than a continuous number
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.05) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 14)
  ) + 
  coord_cartesian(ylim = c(0, 7)) + #Zooming in for 7 second limit for the test
  labs(
    title = "Student Grade Level Deviations from Optimal Response Time",
    x = "Grade Level",
    y = "Distance from Optimal RT (Seconds)"
  )
```

#### SES 
```{r}
ggplot(Study1_X_school %>% 
         filter(phase != "E"),
       aes(x = as.factor(SES10), y = Distance_From_Optimal, fill = as.factor(SES10))) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.05) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 14)
  ) + 
  coord_cartesian(ylim = c(0, 7)) + 
  labs(
    title = "SES Level Deviations from Optimal Response Time",
    x = "SES Level",
    y = "Distance from Optimal RT (Seconds)"
  )
```

#### Nervous Level
```{r}
ggplot(Study1_X_school %>% 
         filter(phase != "E"),
       aes(x = as.factor(Nervous_Level_Num), y = Distance_From_Optimal, fill = as.factor(Nervous_Level_Num))) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.05) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 14)
  ) + 
  coord_cartesian(ylim = c(0, 7)) + 
  labs(
    title = "Nervous Level Deviations from Optimal Response Time",
    x = "Nervous Level",
    y = "Distance from Optimal RT (Seconds)"
  )
```





